The Design and Implementation of the Synchronous Language \CEU

### 2016-09-30 paper.tex
(>) instantaneous dialogue (>) shared-mem-concurrency

Another distinction regarding event handling in comparison to \CEU is that 
Esterel supports same-cycle bi-directional 
communication~\cite{esterel.compiling}, i.e., two threads can react to one 
another during the same cycle due to mutual signal dependencies.
%
\CEU takes a different approach, posing a tradeoff that an \code{await} is only
valid for the next reaction, i.e., if an \code{await} and \code{emit} occur 
simultaneously in parallel trails, the \code{await} does not awake.
%
These \emph{delayed awaits} avoid corner cases of instantaneous termination and 
re-execution of statements in the same reaction (known as \emph{schizophrenic 
statements}~\cite{esterel.constructive}).

@@

Esterel is only deterministic with respect to external behavior: ``the same 
sequence of inputs always produces the same sequence of 
outputs''~\cite{esterel.primer}.
%
However, the execution order for operations within a reaction is 
non-deterministic: ``if there is no control dependency, as in
\code{(call~f1()~||~call~f2())},
the order is unspecified and it would be an error to rely on 
it''~\cite{esterel.primer}.
%
A number of Esterel-inspired synchronous languages enforce an arbitrary 
execution order for statements in multiple lines of execution to achieve 
intra-reaction determinism
(%
\emph{Reactive C}~\cite{rp.rc},                 % 1991
\emph{Protothreads}~\cite{wsn.protothreads},    % 2006
\emph{SOL}~\cite{wsn.sol},                      % 2007
\emph{SC}~\cite{rp.synchc},                     % 2009
and
\emph{PRET-C}~\cite{rp.pretc}%                  % 2010
).
%, e.g., based on lexical order or manual priority assignment.
%
\CEU also takes the deterministic approach and, when multiple trails are active 
during the same reaction, they are scheduled in lexical order, i.e., in the 
order they appear in the program source code.

>>>>

Another distinction regarding event handling in comparison to \CEU is that 
Esterel supports same-cycle bi-directional 
communication~\cite{esterel.compiling}, i.e., two threads can react to one 
another during the same cycle due to mutual signal dependencies
(a.k.a. \emph{instantaneous dialogue}~\cite{rp.book}).
%
\CEU takes a different approach, posing a tradeoff that an \code{await} is only
valid for the next reaction, i.e., if an \code{await} and \code{emit} occur 
simultaneously in parallel trails, the \code{await} does not awake.
%
These \emph{delayed awaits} avoid corner cases of instantaneous termination and 
re-execution of statements in the same reaction (known as \emph{schizophrenic 
statements}~\cite{esterel.constructive}).

@@

Esterel is only deterministic with respect to external behavior: ``the same 
sequence of inputs always produces the same sequence of 
outputs''~\cite{esterel.primer}.
%
However, the execution order for operations within a reaction is 
non-deterministic: ``if there is no control dependency, as in
\code{(call~f1()~||~call~f2())},
the order is unspecified and it would be an error to rely on 
it''~\cite{esterel.primer}.
%
For this reason, Esterel, does not support shared-memory concurrency:
\emph{``if a variable is written by some thread, then it can neither be read
nor be written by concurrent threads''}~\cite{esterel.primer}.
%
A number of Esterel-inspired synchronous languages enforce an arbitrary 
execution order for statements in multiple lines of execution to achieve 
intra-reaction determinism
(%
\emph{Reactive C}~\cite{rp.rc},                 % 1991
\emph{Protothreads}~\cite{wsn.protothreads},    % 2006
\emph{SOL}~\cite{wsn.sol},                      % 2007
\emph{SC}~\cite{rp.synchc},                     % 2009
and
\emph{PRET-C}~\cite{rp.pretc}%                  % 2010
).
%, e.g., based on lexical order or manual priority assignment.
%
\CEU also takes the deterministic approach and, when multiple trails are active 
during the same reaction, they are scheduled in lexical order, i.e., in the 
order they appear in the program source code.
